---
title: "Lab02"
author: "Nicholas Carey"
date: "02/01/2018"
output: html_document
---
###Loading the File
```{r}
load('nba2017-salary-points.RData')
scored <- points1 + points2 + points3 #scored is not a variable that actually exists in this file so I assumed it was the total number of baskets scored and made it myself
ls()
```
###Inspecting the data objects
```{r}
length(player)
length(team)
length(position)
length(salary)
length(scored)
```
All the vectors are the same length
```{r}
anyNA(c(player, team, position, salary, scored, points))
is.vector(player)
is.vector(team)
is.vector(position)
is.vector(salary)
is.vector(scored)
is.vector(points)
head(player)
head(team)
head(position)
head(salary)
head(scored)
head(points)
tail(player)
tail(team)
tail(position)
tail(salary)
tail(scored)
tail(points)
summary(player)
summary(team)
summary(position)
summary(salary)
summary(scored)
summary(points)
```
None of the vectors contain missing values.   
Player and position vectors are **character vectors.**   
Team is a **factor.**    
Salary, points and scored are **numeric vectors** with salary containing *doubles* and scored and points containing *integers*
We know the above information because of the types of elements in each vector and the output from the functions we used.

******

##Subsetting Vectors
```{r}
four <- head(player, n = 4)
four[1]
four[0]
four[-1]
four[-c(1, 2, 3)]
four[5]
four[c(1, 2, 2, 3, 3, 3)]
player[seq(from = 2, to = length(player), by = 2)]
salary[seq(from = 1, to = length(salary), by = 2)]
team[seq(from = 5, to = length(team), by = 5)]
scored[seq(from = 10, to = length(scored), by = 10)]
rev(team[seq(from = 2, to = length(team), by = 2)])
player[team == "GSW" & position == "C"]
player[team == "GSW" | team == "LAL"]
player[position == "PG" | position == "SG" & team == "LAL"]
player[team == "GSW" | team == "LAL" & position == "SF"]
player[which.max(salary)]
player[which.min(salary)]
player[which.max(scored)]
salary[which.max(points)]
max(salary[position == "C"])
team[which.max(scored)]
player[which.max(points3)]
```
###Subsetting with Character Vectors
```{r}
warriors_player <- player[team == "GSW"]
warriors_salary <- salary[team == "GSW"]
warriors_points <- scored[team == "GSW"]
names(warriors_salary) <- warriors_player
warriors_salary["Andre Iguodala"]
warriors_salary[c("Stephen Curry", "Kevin Durant")]
```

******

##Some plotting
```{r}
plot(scored, salary, main = "Points Scored vs Salary")
log_scored <- log(scored)
log_salary <- log(salary)
plot(log_scored, log_salary, main = "Log-Transformed Points Scored vs Salary")
text(log_scored, log_salary, labels = player)
plot(log_scored, log_salary, main = "Log-Transformed Points Scored vs Salary")
text(log_scored, log_salary, labels = abbreviate(player))
plot(warriors_points, warriors_salary, main = "Warriors Scored Points vs Salary")
text(warriors_points, warriors_salary, labels = warriors_player)
log_warriors_points <- log(warriors_points)
log_warriors_salary <- log(warriors_salary)
plot(log_warriors_points, log_warriors_salary, main = "Log-Transformed Warriors Scored Points vs Salary")
text(log_warriors_points, log_warriors_salary, labels = warriors_player)
```

******

##Vectorization and Recycling
```{r}
salary_millions <- salary/1000000
scored_points <- points1 + (2 * points2) + (3 * points3)
sum(scored_points != points) == 0 #checks if scored_points and points are the same
```

******

##Factors
```{r}
is.factor(team)
position_fac <- factor(position)
table(position_fac)
position_fac[team == "GSW"]
position_fac[salary_millions > 15]
table(position_fac[salary_millions > 15])
prop.table(table(team[position == "SG"]))
```

******

##More Plots
```{r}
plot(scored, salary_millions, col = position_fac, main = "Score vs Salary", pch = 10, cex = 0.75, xlab = "Number of Baskets Made", ylab = "Salary in Millions")
legend("bottomright", legend = unique(position_fac), pch = 10, col = unique(position_fac), title = "Postion")
#I had to use the unique() function instead of levels() in the legend so that the legend correctly corresponded  to the plot. Also I used pch = 10 because that is symbol closest in appearance to a basketball
```