---
title: "Lab01"
author: "Nicholas Carey"
date: "1/24/2018"
output: html_document
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Creating Objects
```{r, echo = TRUE}
phone <- 80
transportation <- 20
groceries <- 527
gym <- 10
rent <- 1500
other <- 83
totalExpenses <- phone + transportation + groceries + gym + rent + other
totalExpenses
semesterExpenses <- totalExpenses * 5
semesterExpenses
yearExpenses <- totalExpenses * 10
yearExpenses
```
***
   
###Vectors and Barplots
   
```{r, echo = TRUE}
expenses <- c(phone, transportation, groceries, gym, rent, other)
names(expenses) <- c('phone', 'transport', 'groceries', 'gym', 'rent', 'other')
plotExpense <- barplot(expenses, ylim = c(0,max(expenses) + 0.1 * max(expenses)))
text(plotExpense, expenses, labels = as.character(expenses), pos = 3)
expensesSort <- sort(expenses, decreasing = TRUE)
plotSort <- barplot(expensesSort, ylim = c(0,max(expensesSort) + 0.1 * max(expensesSort)))
text(plotSort, expensesSort, labels = as.character(expensesSort), pos = 3)
```
   
******
   
###Pythagoras Formula
   
$$ c = \sqrt{a^2 + b^2} $$
   
```{r, echo = TRUE}
a <- 3
b <- 4
c <- sqrt(a^2 + b^2)
c
```
   
******

###Binomial Formula
   
####The factorial() function
Probability of flipping a fair coin 5 times and getting heads exactly twice:
```{r, echo = TRUE}
n <- 5
k <- 2
p <- 0.5
nCk <- factorial(n)/(factorial(k)*factorial(n-k)) #n Choose k formula
nCk
probKNP <- nCk * p^k * (1-p)^(n-k) #Binomial probability formula
probKNP
```
   
#### The choose() function
```{r, echo = TRUE}
n <- 5
k <- 2
p <- 0.5
nCk <- choose(n = n, k = k) #choose() function to find n Choose k
nCk
probKNP <- nCk * p^k * (1-p)^(n-k)
probKNP
```
   
Probability of rolling a fair die 10 times and getting exactly 3 sixes:
```{r, echo = TRUE}
n <- 10
k <- 3
p <- 1/6
nCk <- choose(n = n, k = k)
nCk
probKNP <- nCk * p^k * (1-p)^(n-k)
probKNP
```
   
#### The dbinom() function
Probability of 2 heads in 5 coin tosses:
```{r, echo = TRUE}
n <- 5
k <- 2
p <- 0.5
probKNP <- dbinom(x = k, size = n, prob = p)
probKNP
```
    
Probability of 3 sixes in 10 dice rolls:
```{r, echo = TRUE}
n <- 10
k <- 3
p <- 1/6
probKNP <- dbinom(x = k, size = n, prob = p)
probKNP
```

Probability of 2 heads in 5 coin tosses with a biased coin:
```{r, echo = TRUE}
n <- 5
k <- 2
p <- 0.35
probKNP <- dbinom(x = k, size = n, prob = p)
probKNP
```
   
More than 3 heads with a biased coin:
```{r, echo = TRUE}
n <- 5
k <- 3
p <- 0.35
probKNP <- pbinom(q = k, size = n, prob = p, lower.tail = FALSE)
probKNP
```
   
******
   
###Installing Packages
```{r, echo = TRUE, eval = FALSE}
install.packages(c("stringr", "RColorBrewer", "XML"), repos = "http://cran.us.r-project.org")
```
In the code chunk I set `eval = FALSE` to prevent downloading the packages every time it is run.   

******   
    
$$ 3x^2 + 4x + 6 $$
```{r, echo = TRUE}
x <- 2
3*(x^2) + 4*x + 6
```
    
```{r, echo = TRUE}
x <- -3:3
3*(x^2) + 4*x + 6
```
   
Review Questions:   
```{r, echo = TRUE, eval = FALSE}
# example 1
var<-3
Var*2 #output: 6


# example 2
x<-2
2x<-2*x #output: error because variable name starts with a number


# example 3
sqrt4 <- sqrt(4)
sqrt4 #output: 2


# example 4
a number <- 16 #output: error because variable name contains a space


# example 5
"one number" <- 16
`one number` #output: 16
one number   #output: error because this is not a defined object and could not be because it contains a space
```
