---
title: "Lab04"
author: "Nicholas Carey"
date: "2/15/2018"
output: html_document
---
## Abalone
* What's the character delimiter?
    + The character delimiter is ","  
* Is there a row for column names?
    + There is no row for column names.  
* Are there any missing values? If so, how are they codified?
    + There are no missing values
* What is the data type of each column?
    + Sex is a factor. Length, diameter, height, whole weight, shucked weight, and viscera rate are real numbers. Rings is an integer. 
```{r}
url <- "http://archive.ics.uci.edu/ml/machine-learning-databases/abalone/abalone.data"
abalone <- read.table(url, sep = ",")
```
#### Getting a Local Copy of Data
```{r}
origin <- "http://archive.ics.uci.edu/ml/machine-learning-databases/abalone/abalone.data"
destination <- "~/Desktop/Stat133/Lab4/abalone.data"
download.file(origin, destination)
abalone <- read.table(destination, sep = ",")
head(abalone)
tail(abalone)
str(abalone, vec.len = 1)
```
#### Basic Importing
```{r}
# vector of column names
col_names <- c(
    'sex',
    'length',
    'diameter',
    'height',
    'whole_weight',
    'shucked_weight',
    'viscera_weight',
    'shell_weight',
    'rings'
)

# vector of data types (for each column)
col_types <- c(
    'factor',
    'numeric',
    'numeric',
    'numeric',
    'numeric',
    'numeric',
    'numeric',
    'numeric',
    'integer'   
)
abalone <- read.table("~/Desktop/Stat133/Lab4/abalone.data",
                      col.names = col_names,
                      colClasses = col_types,
                      sep = ","
)

#check its structure again
str(abalone, vec.len = 1)
```
#### Your Turn
```{r}
abalone <- read.csv("http://archive.ics.uci.edu/ml/machine-learning-databases/abalone/abalone.data", 
                    col.names = col_names, 
                    colClasses = col_types)
summary(abalone[-1], digits = 4)
sd(abalone$length)
sd(abalone$diameter)
sd(abalone$height)
sd(abalone$whole_weight)
sd(abalone$shucked_weight)
sd(abalone$viscera_weight)
sd(abalone$shell_weight)
sd(abalone$rings)
```

## Pittsburgh Bridges
* Are there column names?  
    + There are no column names in the file.  
* What is the field separator?  
    + The field seperator is ","  
* Are there any missing values?  
    + There are several missing values.  
* What is the character for missing values (if any)?  
    + "?" is used to represent missing value.  
* What is the data type of each variable (i.e. column)?  
    + The data types of the 13 variables are listed in the chunk below in the var_types vector.
```{r}
# download a copy of the data file
origin <- "http://archive.ics.uci.edu/ml/machine-learning-databases/bridges/bridges.data.version1"
destination <- "~/Desktop/Stat133/Lab4/bridges.data"
download.file(origin, destination)

# vector of column names
var_names <- c('ID', 'river', 'location', 'erected_year', 'purpose', 'length', 'lanes', 'Clear-g', 't_or_d', 'material', 'span', 'relative_length', 'type')

# vector of column types
var_types <- c('character', 'factor', 'numeric', 'integer', 'factor', 'integer', 'integer', 'factor', 'factor', 'factor', 'factor', 'factor', 'factor')

# reading the data with read.table()
bridges1 <- read.table( 
  destination, 
  col.names = var_names, 
  colClasses = var_types, 
  sep = ",", 
  na.strings = "?", 
  header = FALSE)

# reading the data with 'read.csv()'
bridges2 <- read.csv(destination, 
                    col.names = var_names, 
                    colClasses = var_types, 
                    na.strings = "?", 
                    header = FALSE)

# setting colClasses to read first 5 columns
bridges3 <- read.table( 
  destination, 
  col.names = var_names, 
  colClasses = c(var_types[1:5], rep('NULL', times = 8)), 
  sep = ",", 
  na.strings = "?", 
  header = FALSE)
bridges3
```
#### Basic Inspection
```{r}
str(bridges1)
summary(bridges1)
head(bridges1)
tail(bridges1)
dim(bridges1)
names(bridges1)
colnames(bridges1)
nrow(bridges1)
ncol(bridges1)
```
#### Optional: Want to do more?
```{r}
old_year <- min(bridges1$erected_year)
oldest <- bridges1$ID[bridges1$erected_year == old_year]
paste(oldest, "is the oldest bridge, built in", old_year)

new_year <- max(bridges1$erected_year)
newest <- bridges1$ID[bridges1$erected_year == new_year]
paste(newest, "is the newest bridge, built in", new_year)

table(bridges1$purpose)
table(bridges1$material)
avrg_length <- round( mean(bridges1$length, na.rm = TRUE), digits = 0)
paste("The average length of the bridges is", avrg_length, "feet.")
plot(x = bridges1$erected_year, 
     y = bridges1$length, 
     main = "Bridges: Year Erected vs. Length",
     xlab = "Year Erected", 
     ylab = "Length",
     col = bridges1$material)
legend("topleft", 
       legend = unique(bridges1$material), 
       col = unique(bridges1$material), 
       title = "Bridge Material",
       pch = 1)
cor_line <- lm(bridges1$length ~ bridges1$erected_year)
abline(cor_line)
cor_eq <- paste("y = ", round(coef(cor_line)[2], digits = 2), "x ", round(coef(cor_line)[1], digits = 0), sep = "")
text(x = 1973, y = 1300, labels = cor_eq)
```

## Creating Data Frames
```{r}
player <- c('Thompson', 'Curry', 'Green', 'Durant', 'Pachulia')
position <- c('SG', 'PG', 'PF', 'SF', 'C')
salary <- c(16663575, 12112359, 15330435, 26540100, 289800)
points <- c(1742, 1999, 776, 1555, 426)
ppg <- c(22.3, 25.3, 10.2, 25.1, 6.1)
rookie <- c(FALSE, FALSE, FALSE, FALSE, FALSE)
nba <- data.frame(player, 
                  position, 
                  salary, 
                  points, 
                  ppg, 
                  rookie)
is.data.frame(nba)
is.list(nba)
nba2 <- data.frame(list(player, 
                        position, 
                        salary, 
                        points, 
                        ppg, 
                        rookie))
nba3 <- data.frame(player, 
                   as.factor(position), 
                   salary, 
                   as.integer(points), 
                   ppg, 
                   rookie, 
                   stringsAsFactors = FALSE)
str(nba3)
nba_bind <- cbind(player, 
                  position, 
                  salary, 
                  points, 
                  ppg, 
                  rookie)
class(nba_bind)
nba_bind <- as.data.frame(nba_bind)
class(nba_bind)
```
## NBA Players Data
```{r}
# download csv file into your working directory
csv <- "https://github.com/ucb-stat133/stat133-fall-2017/raw/master/data/nba2017-players.csv"
download.file(url = csv, 
              destfile = 'nba2017-players.csv')
dat <- read.csv('nba2017-players.csv', 
                stringsAsFactors = FALSE)
```
#### Your Turn
```{r}
# dimensions (# of rows, # of columns)
dim(dat)
# display first few rows
head(dat)
#check the structure
str(dat, vec.len = 1)
#last 5 rows
tail(dat)
summary(dat)
#players shorter than 70 inches
dat[dat$height < 70,]
#names and salaries of centers
dat[dat$position == "C",c("player", "salary")]
durant <- dat[dat$player == "Kevin Durant",]
ucla <- dat[dat$college == "University of California, Los Angeles",]
rookies <- dat[dat$experience == 0,]
rookie_centers <- dat[dat$experience == 0 & dat$position == "C",]
top_players <- dat[dat$games > 50 & dat$minutes > 100,]
#Largest height
max(dat$height)
#Minimum height
min(dat$height)
#Average height
mean(dat$height)
#Tallest player
dat$player[dat$height == max(dat$height)]
#Shortest player
dat$player[dat$height == min(dat$height)]
#Unique NBA teams
unique(dat$team)
#Number of unique NBA teams
length(unique(dat$team))
#Oldest player
dat$player[dat$age == max(dat$age)]
#Median salary of all players
median(dat$salary)
#Median salary of players with 10 or more years experience
median(dat$salary[dat$experience >= 10])
#Median salalry of Shooting Guards nad Point Guards
median(dat$salary[dat$position == "SG" | dat$position == "PG"])
#Median salary of Power Forwards, 29 or more years old and 74 inches or less tall 
median(dat$salary[dat$position == "PF" & dat$age >= 29 & dat$height <= 74])
#Number of players who scored 4 points or fewer
length(dat$player[dat$points <= 4])
#Those bad players
dat$player[dat$points <= 4]
#Player with 0 points
dat$player[dat$points == 0]
#UC Berkeley alumni
length(dat$player[dat$college == "University of California"])
#Notre Dame alumni
ND <- dat$player[dat$college == "University of Notre Dame"]
ND
#Number of ND alumni
length(ND)
#Heavy players
heavy <- dat$player[dat$weight > 260]
length(heavy)
heavy
#Number of players not from US colleges
length(dat$player[dat$college == ""])
#Player with highest points per minute
dat$player[which.max(dat$points/dat$minutes)]
#Player with highest 3-pointers per minute
dat$player[which.max(dat$points3/dat$minutes)]
#Player wiht highest 2-pointers per minute
dat$player[which.max(dat$points2/dat$minutes)]
#Player with highest free throws per minute
dat$player[which.max(dat$points1/dat$minutes)]
gsw <- dat[dat$team == "GSW", c("player", "height", "weight")]
gsw[order(gsw$height),]
gsw[order(gsw$weight, decreasing = TRUE),]
#Top 5 Paid
head(dat[order(dat$salary, decreasing = TRUE), c("player", "team", "salary")], n = 5)
#Top 10 3-point shooters
head(dat[order(dat$points3, decreasing = TRUE), c("player", "team", "points3")], n = 10)
```
#### Group By, Your Turn
```{r}
#Average height, weight and age by position
aggregate(dat[, c("height", "weight", "age")], 
          by = list(dat$position), 
          FUN = mean)
#Average height, weight, and age by team
aggregate(dat[, c("height", "weight", "age")], 
          by = list(dat$team), 
          FUN = mean)
#Average height, weight, and age by team and position
aggregate(dat[, c("height", "weight", "age")], 
          by = list(dat$team, dat$position), 
          FUN = mean)
#Median min, median, mean, and max salary by team and position
aggregate(dat$salary, 
          by = list(dat$team, dat$position), 
          FUN = function(salary) {
            c(min = min(salary), median = median(salary), mean = mean(salary), max = max(salary))
            }
          )
```